A dynamic programing algorithm to assess the reliability of inferences from network enrichment analysis

### 2018-04-30 doc/genesetdp.tex
x

Gene enrichment analysis (GEA) is commonly used to infer function from sets of analytes such as genes, transcripts proteins or metabolites\cite{tavazoie1999systematic,khatri2012ten}. The technique is used to estimate which functional modules such as complexes or pathways that diverge between different biological conditions, such as disease state or treatment group. Particularly in expression analysis, GEA is used to assess alternation in activity of functional models by examining significantly different concentrations of analytes between experimental conditions.

@@

Here, we present a dynamic programing algorithm, that we dubbed GeneSetDP, to calculate the full score distribution of any query of a given number of analytes. The algorithm enables us to calculate well defined and exact statistics of inferences from NEA.  Using simulations we demonstrated that the algorithm produced unbiased statistics.

>>>>

Gene enrichment analysis (GEA) is commonly used to infer function from sets of analytes such as genes, transcripts proteins or metabolites\cite{tavazoie1999systematic,khatri2012ten}. The technique predict which functional modules, such as complexes or pathways, that are overrepresentated among a set of identified analytes. One prominent application of the technique is expression analysis, where GEA is regularly used to assess alternation in pathway activity by examining significantly different concentrations of analytes between biological conditions, such as disease state or treatment group.

@@

Here, we present a dynamic programing algorithm, that we dubbed GeneSetDP, that calculate the full score distribution of any query of a given size. The algorithm enables us to calculate well defined and exact statistics of inferences from NEA.  Using simulations we demonstrated that the algorithm produced unbiased statistics.
